\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2022-01-13 papers/ast22/main.tex
Replaced FIFO -> FIFO. Updated replication package link.

%% Heuristics to improve testing loop
Additionally, the testing loop of \mutagen incorporates two novel heuristics
that help finding bugs faster and more reliably (\S \ref{sec:heuristics}).
%
In the first place, our tool uses first-in first-out (FIFO) scheduling with
priority to enqueue interesting test cases for mutation.
%
This scheduling algorithm is indexed by the novelty of each new test case with
respect to the ones already seen.
%
In this light, interesting test cases that discover new parts of the code
earlier during execution are given a higher priority.
%
Moreover, our scheduling allows the testing loop to jump back and forth between
mutable test cases as soon as a new more interesting one is enqueued for
mutation, eliminating potential delays when the mutation queues grow large.

@@

\subsection{Priority FIFO Scheduling}

@@

\caption{\label{algo:mutagen:fifo}Priority FIFO Heuristic}
\end{algorithm}

@@

Using this mechanism, we can modify \mutagen's base testing loop by replacing
each mutation queue with a priority queue indexed by the branching depth of each
new execution.
%
These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
%
Statements in {\color{red} red} indicate important changes to the base
algorithm, whereas ellipses denote parts of the code that are not relevant for
the point being made.

@@

In both case studies, we show how the two \mutagen heuristics described in \S
\ref{sec:heuristics} affect the testing performance by individually disabling
them.
%
We call these variants \textit{no FIFO} and \textit{no reset}, respectively.
%
In the case of \textit{no reset}, the amount of random mutations is no longer
controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
throughout the execution of testing loop.

@@

In terms of the \mutagen heuristics,
%
%% we can also arrive to some conclusions.
%% %
%% Firstly,
%
this case study is not strongly affected by using FIFO scheduling.
%% --- the same
%% benchmark takes only roughly 20\% longer to complete when disabling this
%% heuristic.
%
The reason behind this is that the mutation queues remain empty most of the time
(generation mode), and when a new interesting candidate gets inserted, all its
mutants (and their descendants) are processed before the next one is enqueued.
%
Hence, the small proportion of time this heuristic is active tends to be
inconsequential.
%% under this case study.

@@

In terms of the \mutagen heuristics, we observe two clear phenomena.
%
On one hand, disabling the FIFO scheduling (case \textit{no FIFO}), the
hardest-to-find bugs (\#7, \#10, and \#1) are no longer found on every run.
%
Moreover, although bugs \#2, \#9, \#8, and \#6 are found with 100\% success rate
across runs, it takes several times longer for \mutagen to find them.
%
On the other hand, disabling the trace saturation heuristic (case \textit{no
  reset}) and using a fixed number of 25 random mutation does not affect the
effectiveness of our tool, apart from adding again a constant time overhead.
%% ---
%% suggesting that the value we arbitrarily chose for this parameter is too large
%% and should be left to be handled automatically by our tool.

If we now pay attention to the bugs injected into the interpreter (Fig.
\ref{fig:results:wasm} right), we notice that finding bugs requires
substantially more time (minutes instead of seconds), since both interpreters
need to validate and run the inputs before producing a result to compare.
%
We can also observe a significant improvement in the performance of \quickcheck
in terms of failure rate.
%
This is of no surprise: we deliberately reduced the search space by using a
module stub when defining \texttt{prop\_interpreter} in our test suite.
%
Notably, \quickcheck finds counterexamples for the bug \#14 almost instantly.
%
The reason behind this is that this bug can be found using a very small
counterexample, and \quickcheck prefers sampling small test cases at the
beginning of the testing loop.
%
Our tool follows this same approach when in generation mode.
%
However, when a test case is found interesting, the scheduler does not take its
size into account while computing its priority --- future work should
investigate this possibility.
%
Nonetheless, \mutagen still outperforms \quickcheck on the remaining bugs both
in terms of failure rate and mean time to failure.
%
Moreover, by disabling the FIFO scheduling we observe that the performance of
our tool becomes unstable, where bugs \#12 and \#11 take substantially less time
to be found, whereas bugs \#15 and \#14 cannot be found on every run.
%
Finding where the best trade-off lies in cases where the heuristics cause mixed
results like this is another interesting direction to aim our future work.

>>>>

%% Heuristics to improve testing loop
Additionally, the testing loop of \mutagen incorporates two novel heuristics
that help finding bugs faster and more reliably (\S \ref{sec:heuristics}).
%
In the first place, our tool uses last-in first-out (LIFO) scheduling with
priority to enqueue interesting test cases for mutation.
%
This scheduling algorithm is indexed by the novelty of each new test case with
respect to the ones already seen.
%
In this light, interesting test cases that discover new parts of the code
earlier during execution are given a higher priority.
%
Moreover, our scheduling allows the testing loop to jump back and forth between
mutable test cases as soon as a new more interesting one is enqueued for
mutation, eliminating potential delays when the mutation queues grow large.

@@

\subsection{Priority LIFO Scheduling}

@@

\caption{\label{algo:mutagen:lifo}Priority LIFO Heuristic}
\end{algorithm}

@@

Using this mechanism, we can modify \mutagen's base testing loop by replacing
each mutation queue with a priority queue indexed by the branching depth of each
new execution.
%
These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
%
Statements in {\color{red} red} indicate important changes to the base
algorithm, whereas ellipses denote parts of the code that are not relevant for
the point being made.

@@

In both case studies, we show how the two \mutagen heuristics described in \S
\ref{sec:heuristics} affect the testing performance by individually disabling
them.
%
We call these variants \textit{no LIFO} and \textit{no reset}, respectively.
%
In the case of \textit{no reset}, the amount of random mutations is no longer
controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
throughout the execution of testing loop.

@@

In terms of the \mutagen heuristics,
%
%% we can also arrive to some conclusions.
%% %
%% Firstly,
%
this case study is not strongly affected by using LIFO scheduling.
%% --- the same
%% benchmark takes only roughly 20\% longer to complete when disabling this
%% heuristic.
%
The reason behind this is that the mutation queues remain empty most of the time
(generation mode), and when a new interesting candidate gets inserted, all its
mutants (and their descendants) are processed before the next one is enqueued.
%
Hence, the small proportion of time this heuristic is active tends to be
inconsequential.
%% under this case study.

@@

In terms of the \mutagen heuristics, we observe two clear phenomena.
%
On one hand, disabling the LIFO scheduling (case \textit{no LIFO}), the
hardest-to-find bugs (\#7, \#10, and \#1) are no longer found on every run.
%
Moreover, although bugs \#2, \#9, \#8, and \#6 are found with 100\% success rate
across runs, it takes several times longer for \mutagen to find them.
%
On the other hand, disabling the trace saturation heuristic (case \textit{no
  reset}) and using a fixed number of 25 random mutation does not affect the
effectiveness of our tool, apart from adding again a constant time overhead.
%% ---
%% suggesting that the value we arbitrarily chose for this parameter is too large
%% and should be left to be handled automatically by our tool.

If we now pay attention to the bugs injected into the interpreter (Fig.
\ref{fig:results:wasm} right), we notice that finding bugs requires
substantially more time (minutes instead of seconds), since both interpreters
need to validate and run the inputs before producing a result to compare.
%
We can also observe a significant improvement in the performance of \quickcheck
in terms of failure rate.
%
This is of no surprise: we deliberately reduced the search space by using a
module stub when defining \texttt{prop\_interpreter} in our test suite.
%
Notably, \quickcheck finds counterexamples for the bug \#14 almost instantly.
%
The reason behind this is that this bug can be found using a very small
counterexample, and \quickcheck prefers sampling small test cases at the
beginning of the testing loop.
%
Our tool follows this same approach when in generation mode.
%
However, when a test case is found interesting, the scheduler does not take its
size into account while computing its priority --- future work should
investigate this possibility.
%
Nonetheless, \mutagen still outperforms \quickcheck on the remaining bugs both
in terms of failure rate and mean time to failure.
%
Moreover, by disabling the LIFO scheduling we observe that the performance of
our tool becomes unstable, where bugs \#12 and \#11 take substantially less time
to be found, whereas bugs \#15 and \#14 cannot be found on every run.
%
Finding where the best trade-off lies in cases where the heuristics cause mixed
results like this is another interesting direction to aim our future work.
